Water, High-Altitude Condensates, and Methane Depletion in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}

### 2017-08-14 ms.tex
conclu

\begin{figure*}
\includegraphics[width = \textwidth]{Figures/spectrum.pdf}
\caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals). The features at $1.15$ and $1.4$ $\mu$m are due to water absorption. The right-hand axis indicates the normalized atmospheric scale height assuming a $100\times$ solar composition.}
\label{fig:spectrum}
\end{figure*}

@@

%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.  
\section{Conclusions} \label{sec:discuss}
We presented an analysis of the atmospheric composition of WASP-107b based on its measured near-infrared transmission spectrum and interior structure modeling of the planet's mass and radius.  Key results from this analysis include:

\begin{itemize}
\item{The upper limit on atmospheric metallicity is $30\times$ solar (assuming the star is young) or FIXME solar for an older system.  Interior structure modeling shows that higher metallicity envelopes would decrease the radius below the observed value.  Based on the Solar System trend of increasing planet metallicity with decreasing planet mass, WASP-107b is expected to have a metallicity of $30\times$ solar, just at the edge of our allowed range \citep{kreidberg14b}. Further characterization of the stellar age is needed to determine whether WASP-107b is consistent with expectations, or discrepant with the Solar System trend, which would suggest an intriguing difference in its formation history.}
\item{The methane abundance is depleted relative to expectations for a solar abundance pattern. This could be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.   Another possibility is that the planet has a hot interior.}
\item{Optically thick condensates are present at high altitudes ($0.1 - 3$ mbar).  Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observed . Clouds form too deep in the atmosphere }
\end{itemize}

%Some directly imaged planets in this temperature regime are also unexpectedly depleted in methane when compared to similar temperature brown dwarfs \citep{skemer14}.   The leading hypothesis of this depletion \citep{zahnle14} is due to the CH4-CO quench point occurring in the CO stability field due to the low gravity of self-luminous planets when compared to brown dwarfs.

%In order this mechanism to work for neptune massed planet, the internal temperature would have to be near ~500K to get the deep temperatures high enough to thermochemically deplete methane at the quench point. Given the age of the star, it is unlikely that such high internal temperatures are still due to heat of formation. It is possible that tidal heating could increase the internal temperature (Agundez et al. 2014; Morley et al. 2017) but that would require a negligible eccentricity, of which is currently uncertain for WASP107b.

@@

Combine data from Spitzer/HST
JWST GTO

>>>>

\begin{figure*}
\includegraphics[width = \textwidth]{Figures/spectrum.pdf}
\caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals). The features at $1.15$ and $1.4$ $\mu$m are due to water absorption. The right-hand axis indicates the normalized atmospheric scale height assuming a $10\times$ solar composition.}
\label{fig:spectrum}
\end{figure*}

@@

%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.  
\section{Conclusions} \label{sec:discuss}
We analyzed the atmospheric composition of WASP-107b based on its measured near-infrared transmission spectrum and interior structure models of the planet's mass and radius.  Key results from this analysis include:

\begin{itemize}
\item{The upper limit on atmospheric metallicity is $30\times$ solar (assuming the star is young) or FIXME solar for an older system.  Higher metallicity envelopes are not allowed because they decrease the planet's radius below the observed value.  Based on the Solar System trend of increasing planet metallicity with decreasing planet mass, WASP-107b is expected to have a metallicity of $30\times$ solar, just at the edge of our allowed range \citep{kreidberg14b}. Further characterization of the stellar age is needed to determine whether WASP-107b is consistent with expectations, or discrepant with the Solar System trend, which would suggest an intriguing difference in its formation history. Either way, this planet stands in contrast to the similar mass and temperature planet GJ 436b, which is thought to have a metallicity greater than $100\times$ solar \citep{morley17}. Together, these exoplanets suggest a greater diversity of compositions than exists in the Solar System.}
\item{The methane abundance is depleted relative to expectations for a solar abundance pattern. This may be due to an instrinsically low carbon-to-oxygen ratio, which could arise from accretion of water-rich planetesimals \citep{mordasini16, espinoza17}.  Another possibility is that the planet has a hot interior ($\sim500$ K), and abundances are quenched at pressure levels where CO is stable \citep[as observed in some directly imaged planets;][]{skemer14, zahnle14}. Such a high internal temperature could be due to latent heat of formation if the planet is very young, and/or tidal heating \citep{fortney08, morley17}.} 
\item{Optically thick condensates are present at high altitudes ($0.1 - 3$ mbar).  Existing models of physically-motivated cloud and haze formation do not satisfactorily reproduce the observed . Clouds form too deep in the atmosphere . \citep{crossfield17}} GJ 1214b
\end{itemize}

@@
